---
title: "PhIrDO_MicroelectrodProfiles"
output: html_document
date: "2023-08-09"
editor_options: 
  chunk_output_type: console
---

```{r}
library(stringr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tidyr) 
library(hms)
```

#Data
##Microelectrode Profiles
```{r}
#Each Raw data file corresponds to one sequence programmed and run
rawDO1<-read.table("Microelectrode profile data/28072023002440_TFS_PhIr_5cmdepth_2mm_365/28072023002440_data.txt", sep="\t", header = TRUE)
rawDO2<-read.table("Microelectrode profile data/28072023003700_TFS_PhIr_5cm_1hr_diurnal_368/28072023003700_data.txt", sep="\t", header = TRUE)
rawDO3<-read.table("Microelectrode profile data/28072023142101_TFS_PhIr_5cm_1hr_diurnal_370/28072023142101_data.txt", sep="\t", header = TRUE)
rawDO4<-read.table("Microelectrode profile data/29072023131024_TFS_PhIr_5cm_1hr_diurnal_371/29072023131024_data.txt", sep="\t", header = TRUE)
rawDO5<-read.table("Microelectrode profile data/29072023213519_TFS_PhIr_5cm_30min_2d_374/29072023213519_data.txt", sep="\t", header = TRUE)

#Adding the file name for each set of raw data in the "file" column
rawDO1$file<-"28072023002440_data.txt"
rawDO2$file<-"28072023003700_data.txt"
rawDO3$file<-"28072023142101_data.txt"
rawDO4$file<-"29072023131024_data.txt"
rawDO5$file<-"29072023213519_data.txt"

#Adding the program name corresponding to each sequence run in the "program" column
rawDO1$program<-"28072023002440_TFS_PhIr_5cmdepth_2mm_365"
rawDO2$program<-"28072023003700_TFS_PhIr_5cm_1hr_diurnal_368"
rawDO3$program<-"28072023142101_TFS_PhIr_5cm_1hr_diurnal_370"
rawDO4$program<-"29072023131024_TFS_PhIr_5cm_1hr_diurnal_371"
rawDO5$program<-"29072023213519_TFS_PhIr_5cm_30min_2d_374"

#merging the five raw data files together using rbind into a new data frame called microDOraw
microDOraw<-rbind(rawDO1, rawDO2)
microDOraw<-rbind(microDOraw, rawDO3)
microDOraw<-rbind(microDOraw, rawDO4)
microDOraw<-rbind(microDOraw, rawDO5)

#Keeping only relevant columns
microDOraw<-microDOraw %>%
  select(Time, Channel.1..mV., Channel.2..mV., Channel.1_cal..µmol.L., Channel.2_cal..µmol.L., Motor.1..mm., Motor.2..mm., GroupId, file, program)

#making program a factor
microDO$program<-as.factor(microDO$program)

#renaming columns to get rid of annoying extra punctuations
names(microDOraw)[names(microDOraw)=="Channel.1..mV."]<-"Channel1_mV"
names(microDOraw)[names(microDOraw)=="Channel.2..mV."]<-"Channel2_mV"
names(microDOraw)[names(microDOraw)=="Channel.1_cal..µmol.L."]<-"Channel1_DO_mgL"
names(microDOraw)[names(microDOraw)=="Channel.2_cal..µmol.L."]<-"Channel2_DO_mgL"
names(microDOraw)[names(microDOraw)=="Motor.1..mm."]<-"Motor1.mm"
names(microDOraw)[names(microDOraw)=="Motor.2..mm."]<-"Motor2.mm"

#the Time column contains date and time. Code below splits into a Date and Time column. new column with only time is named Time2
microDOraw[c("Date", "Time2")]<-str_split_fixed(microDOraw$Time, " ", 2)

#telling R to format the date as a date in a new column called RDate
microDOraw$RDate<-as.POSIXct(microDOraw$Date, format="%d%m%Y")

#telling R to format original Time column as a date time in eastern daylight time in new column called RDateTimeEDT
microDOraw$RDateTimeEDT<-as.POSIXct(microDOraw$Time, format="%d%m%Y %H%M%S")

#Converting time to AlaskaTime in new column called RDateTimeAST
microDOraw$RDateTimeAST<-as.POSIXct(microDOraw$RDateTimeEDT-hours(4))

#gathering data from the 2 DO sensors into a a single column, creating a new data frame called micro DO to use for most data processing and analysis from here on out.

microDO<-microDOraw %>%
  #selecting only the relevant columns
  select(Time, Channel1_DO_mgL, Channel2_DO_mgL, Motor1.mm, Motor2.mm, program, RDateTimeEDT, RDateTimeAST) %>%
  #gathering the 2 DO sensors into a single column
  gather(key="DOsensor", value="DO.mgL", Channel1_DO_mgL:Channel2_DO_mgL)
#making a column of just times using the as_hms() function in the hms package
microDO$RTime<-as_hms(microDO$RDateTimeAST)
#making a column of just date for faceting by date in graphs
microDO$RDate<-as_date(microDO$RDateTimeAST)

#Make a column that's the hour w/o minutes and seconds that a profile was taken, which will help with graphing later

#Found example here: https://stackoverflow.com/questions/16444242/rounding-times-to-the-nearest-hour-in-r

microDO$RHour<-as.numeric(format(round(microDO$RDateTimeAST, units="hours"), format="%H"))
#Hour is just a number, not a date time or time object.
plot(microDO$RDateTimeAST~microDO$RHour)
```

###Notes on data
28072023002440_TFS_PhIr_5cmdepth_2mm_365 is a trial run, do not use
28072023003700_TFS_PhIr_5cm_1hr_diurnal_368, 28072023003700_data_txt contains data for profiles that got from Motor 1=0 to 49.969 and elapse for 895 seconds and occur every 9,123 seconds (~2.5 hours)
Laptop was set to Eastern Time Zone. To convert to local alaska time, subtract 4 hours
Programs:
[1] "28072023002440_TFS_PhIr_5cmdepth_2mm_365"    "28072023003700_TFS_PhIr_5cm_1hr_diurnal_368"
[3] "28072023142101_TFS_PhIr_5cm_1hr_diurnal_370" "29072023131024_TFS_PhIr_5cm_1hr_diurnal_371"
[5] "29072023213519_TFS_PhIr_5cm_30min_2d_374"  

"28072023002440_TFS_PhIr_5cmdepth_2mm_365" is a 
Single profile, depth recorded on Motor2.mm

###Omitting data from first program from data being used

```{r}
microDO<-microDO%>%
  filter(program!="28072023002440_TFS_PhIr_5cmdepth_2mm_365")
```
###Giving each profile a unique name
```{r}

#Each profile in all four of the programs has 26 points, which makes things easier

#https://stackoverflow.com/questions/32078578/how-to-group-by-every-7-rows-and-aggregate-those-7-values-by-median

#theData$seven_day_index <- c(0, rep(1:(nrow(theData)-1)%/%7))

microDO$ProfIndex <- c(0, rep(1:(nrow(microDO)-1)%/%26) )
#I don't know how this works but it does!

##Name each profile with the date and time stamp of the first data point

#https://stackoverflow.com/questions/69375473/r-how-to-replace-all-rows-in-a-group-with-a-value-from-first-row-in-a-df

#First() function in dplyr gets the first value in each, then adding in the sensor name
microDO<-microDO %>%
  group_by(ProfIndex) %>%
  mutate(ProfileID = paste(first(Time), DOsensor))

#also make one with just time time for co-graphing profiles taken at the same time
microDO<-microDO %>%
  group_by(ProfIndex) %>%
  mutate(ProfileIDTime = first(Time) )

#and make a new column that is the start Date Time
microDO<-microDO %>%
  group_by(ProfIndex) %>%
  mutate(ProfileStart = first(as.POSIXct(RDateTimeAST)) )

#making a new column with just the Time of start Date Time
microDO$RProfileStartTime<-as_hms(microDO$ProfileStart)
```
##Presens FibOx DO Data
```{r}
#Reading in Dissolved oxygen concentrations measured with hand-held Presens FibOx cable
FiboxDOraw<-read.csv("PhIr2023_DissolvedOxygenMeasurements.csv")

#Removing ghost rows by omitting anything with DO is na becuase every row should have a DO value
FiboxDOraw<-FiboxDOraw %>%
  filter(!is.na(DO.mgL))

#where were these data collected? location coordinates are in FiboxDOLoc.csv
##Added coordinates manually to FiboxDOLoc.csv.  Coordinates in FiboxDOLoc.csv are approximate. For more specific locations, see hand-drawn maps in field notes, which are also annotated in 2023July_DOCampaign_Notes.pptx . More accurate coordinates could be determined using the 2021 RTK coordinates, high-resolution georeferenced aerial imagery than what’s available on Google Earth, and field notes maps.

##Getting Dates & Times in order
#Making a DateTime field 
FiboxDOraw$RDate<-as.POSIXct(FiboxDOraw$Date, format="%d-%b-%y")
#Pasting Date and Time together to make a DateTime column
FiboxDOraw$DateTime<-paste(FiboxDOraw$RDate, FiboxDO$Time)
#telling R that it's a Date Tiem using as.POSIXct
FiboxDOraw$RDateTime<- as.POSIXct (FiboxDOraw$DateTime, format="%Y-%m-%d %H:%M")

#omitting data point from July 31 with unclear depth note in field notes (under depth says "multiplex..." for some reason)

FiboxDO<-FiboxDOraw %>%
  filter(!is.na(as.numeric(Depth.cm)))

##Telling R which ones should be numeric that aren't
FiboxDO$Depth.cm<-as.numeric(FiboxDO$Depth.cm)
FiboxDO$Temp.degC<-as.numeric(FiboxDO$Temp.degC)

##Making an Area Site column

FiboxDO$AreaSite<-paste(FiboxDO$Area,FiboxDO$Site)


##Trying to make a column with only time of day
# starting with this as a resource 
##https://r4ds.hadley.nz/datetimes.html
##the above website says: "In this chapter we are going to focus on dates and date-times as R doesn’t have a native class for storing times. If you need one, you can use the hms package."
##So, turning to the hms package: https://hms.tidyverse.org/

library(hms)
FiboxDO$Time<-as_hms(FiboxDO$RDateTime)
#Testing to see if it work and how it graphs
ggplot(FiboxDO, aes(x=Time, y=DO.mgL))+
  geom_point()
#Looks like it worked! yay.
``` 
#Summary Stats
##FiboxDO data
```{r}
##Summarizing DO in top 10 cm of surface water
FiboxDO.surf<-FiboxDO %>%
  filter(Depth.cm < 10) 
summary(FiboxDO.surf)
```


#Graphs
##Microelectrode DO Profiles
###Graphs of DO vs. time, color coded by depth, shape coded by sensor

```{r}
#all of the data except the first program
microDO %>%
  filter(program!= "28072023002440_TFS_PhIr_5cmdepth_2mm_365" )%>%
ggplot( aes(x=RDateTimeAST, y=DO.mgL, color=Motor1.mm, shape=program))+
  geom_point()+
  scale_color_gradient(low="blue", high="orange")

```

###Graphs of DO vs. Depth--traditional profiles

```{r}
ggplot(microDO, aes(x=DO.mgL, y=Motor1.mm, color=as.POSIXct(RDateTimeAST), shape=DOsensor))+
  geom_point()+
  scale_y_reverse()
```

####Program 1, faceted by profile start time
```{r}
microDO%>%
  filter(program=="28072023003700_TFS_PhIr_5cm_1hr_diurnal_368") %>%
  ggplot(aes(x=DO.mgL, y=Motor1.mm, shape=DOsensor))+
  geom_point()+
  scale_y_reverse()+
  facet_wrap(~ProfileStart)
```

####Program 2, faceted by profile start time
```{r}
microDO%>%
  filter(program=="28072023142101_TFS_PhIr_5cm_1hr_diurnal_370") %>%
  ggplot(aes(x=DO.mgL, y=Motor1.mm, shape=DOsensor))+
  geom_point()+
  scale_y_reverse()+
  facet_wrap(~ProfileStart)
```

####Program 3, faceted by profile start time
```{r}
microDO%>%
  filter(program=="29072023131024_TFS_PhIr_5cm_1hr_diurnal_371") %>%
  ggplot(aes(x=DO.mgL, y=Motor1.mm, shape=DOsensor))+
  geom_point()+
  scale_y_reverse()+
  facet_wrap(~ProfileStart)
```

####Program 4, faceted by profile start time
this one has so many profiles, splitting it into sections by date
```{r}
##getting a list of all the profile start date times
microDO4<-microDO%>%
  filter(program== "29072023213519_TFS_PhIr_5cm_30min_2d_374")
unique(microDO4$ProfileStart)
length(unique(microDO4$ProfileStart))

#plotting profiles in groups of 16

microDO%>%
  filter(program== "29072023213519_TFS_PhIr_5cm_30min_2d_374") %>%
  filter(RDateTimeAST<as.POSIXct("2023-07-30 01:50")) %>%
  ggplot(aes(x=DO.mgL, y=Motor1.mm, shape=DOsensor))+
  geom_point()+
  scale_y_reverse()+
  facet_wrap(~ProfileStart)

microDO%>%
  filter(program== "29072023213519_TFS_PhIr_5cm_30min_2d_374") %>%
  filter(RDateTimeAST>as.POSIXct("2023-07-30 01:50") & RDateTimeAST<as.POSIXct("2023-07-30 10:20" )) %>%
  ggplot(aes(x=DO.mgL, y=Motor1.mm, shape=DOsensor))+
  geom_point()+
  scale_y_reverse()+
  facet_wrap(~ProfileStart)

microDO%>%
  filter(program== "29072023213519_TFS_PhIr_5cm_30min_2d_374") %>%
  filter(RDateTimeAST>as.POSIXct("2023-07-30 10:20") & RDateTimeAST<as.POSIXct("2023-07-30 18:40" )) %>%
  ggplot(aes(x=DO.mgL, y=Motor1.mm, shape=DOsensor))+
  geom_point()+
  scale_y_reverse()+
  facet_wrap(~ProfileStart)

##Looks like the motor goofed after the 2023-7-30 13:01 profiles 

microDO%>%
  filter(program== "29072023213519_TFS_PhIr_5cm_30min_2d_374") %>%
  filter(RDateTimeAST>as.POSIXct("2023-07-30 18:40") & RDateTimeAST<as.POSIXct("2023-07-31 3:10" )) %>%
  ggplot(aes(x=DO.mgL, y=Motor1.mm, shape=DOsensor))+
  geom_point()+
  scale_y_reverse()+
  facet_wrap(~ProfileStart)

microDO%>%
  filter(program== "29072023213519_TFS_PhIr_5cm_30min_2d_374") %>%
  filter(RDateTimeAST>as.POSIXct("2023-07-31 3:10")) %>%
  ggplot(aes(x=DO.mgL, y=Motor1.mm, shape=DOsensor))+
  geom_point()+
  scale_y_reverse()+
  facet_wrap(~ProfileStart)



```
###Graphs of DO vs. Depth, faceted by date, color coded by time
```{r}
ggplot(microDO, aes(x=DO.mgL, y=Motor1.mm, color=as.POSIXct(RTime), shape=DOsensor))+
  geom_point()+
  scale_y_reverse()+
  facet_wrap(~RDate)
```
####Individual Dates, faceted by profile, labeled with start time
```{r}
microDO %>%
  filter(RDate==as.POSIXct("2023-07-29"))%>%
  ggplot(aes(x=DO.mgL, y=Motor1.mm))+
  geom_point()+
  scale_y_reverse() +
  facet_wrap(~RProfileStartTime)

##Next step: make a column rounding to the start hour so I can make a grid of day by time. Later, chnage the label of each plot to the actual start time.


ggplot(microDO, aes(x=DO.mgL, y=Motor1.mm))+
  geom_point(size=0.1)+
  scale_y_reverse() +
  facet_grid(rows=vars(RDate), cols=vars(RHour))
ggsave("AllDOProfiles.png", width=10, height=3, units="in")
```

##Presens Fibox DO profiles

###All the data color coded by Date Time
```{r}
ggplot(FiboxDO, aes(x=DO.mgL, y=Depth.cm, color=RDateTime, shape=AreaSite))+
  geom_point()+
  scale_y_reverse()

```
###All the data color coded by Time
```{r}
ggplot(FiboxDO, aes(x=DO.mgL, y=Depth.cm, color=Time, shape=AreaSite))+
  geom_point()+
  scale_y_reverse()
#why isn't there a legend for the color-coded time? also looks like it's reading time as categorical

```
